---
title: Word Rank Slope Charts
date: '2020-03-17'
categories: [tidytext, ggplot2]
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(
  echo = TRUE,
  cache = TRUE,
  fig.width = 7, 
  fig.align = 'center',
  fig.asp = 0.618, # 1 / phi
  out.width = "700px")
knit_hooks$set(optipng = hook_optipng)
opts_chunk$set("optipng" = "-o5")
```

I have been working on visualizing how different kinds of words are used in texts and I finally found a good visualization style with the [slope chart](https://datavizproject.com/data-type/slope-chart/).
More specifically I'm thinking of two groups of paired words.

## Packages `r emo::ji("package")`

```{r, message=FALSE}
library(tidyverse)
library(hcandersenr)
library(tidytext)
library(paletteer)
library(ggrepel)
```

## Minimal Example `r emo::ji("keycap_1")`

First I'll walk you through a minimal example of how the chart is created.
Afterward, I have created a function to automate the whole procedure so we can quickly iterate.
We start by an example of gendered words in fairly tales by H.C. Andersen using the [hcandersenr](https://github.com/EmilHvitfeldt/hcandersenr) package.
We start by generating a data.frame of paired words.
This is easily done using the `tribble()` function.

```{r}
gender_words <- tribble(
  ~men, ~women,
  "he", "she",
  "his", "her",
  "man", "woman",
  "men", "women",
  "boy", "girl",
  "he's", "she's",
  "he'd", "she'd",
  "he'll", "she'll",
  "himself", "herself"
)
```

Next, we are going to tokenize and count the tokens in the corpus,

```{r}
ordered_words <- hcandersen_en %>% 
  unnest_tokens(word, text) %>% 
  count(word, sort = TRUE) %>% 
  pull(word)
```

Next, we are going to get the index for each word, which we will put on a log scale since it will be [easier to visualize](https://en.wikipedia.org/wiki/Zipf%27s_law).
Next, we will calculate a slope between the points and add the correct labels.

```{r}
gender_words_plot <- gender_words %>%
  mutate(male_index = match(men, ordered_words),
         female_index = match(women, ordered_words)) %>%
  mutate(slope = log10(male_index) - log10(female_index)) %>%
  pivot_longer(male_index:female_index) %>%
  mutate(value = log10(value),
         label = ifelse(name == "male_index", men, women)) %>%
  mutate(name = factor(name, c("male_index", "female_index"), c("men", "women")))
```

Next, we are going to manually calculate the limits to make sure a diverging color scale will have the colors [done directly](https://www.hvitfeldt.me/blog/center-continuous-palettes-in-ggplot2/).

```{r}
limit <- max(abs(gender_words_plot$slope)) * c(-1, 1)
```

Lastly, we just put everything into ggplot2 and voila!!

```{r}
gender_words_plot %>%
  ggplot(aes(name, value, group = women, label = label)) +
  geom_line(aes(color = slope)) +
  scale_y_reverse(labels = function(x) 10 ^ x) +
  geom_text() +
  guides(color = "none") +
  scale_color_distiller(type = "div", limit = limit) +
  theme_minimal() +
  theme(panel.border = element_blank(), panel.grid.major.x = element_blank()) +
  labs(x = NULL, y = "Word Rank") +
  labs(title = "Masculine gendered words appeared more often in H.C. Andersen's fairy tales")
```

## Make it into a function `r emo::ji("sparkles")`

This function is mostly the same as the code you saw earlier.
Main difference is using `.data` from [rlang](https://rlang.r-lib.org/reference/tidyeval-data.html) to generalize.
The function also includes other beautifications such as improved themes and theme support with [paletteer](https://github.com/EmilHvitfeldt/paletteer).

```{r}
plot_fun <- function(words, ref, palette = "scico::roma", ...) {
  
  names <- colnames(ref)
  
  ordered_words <- names(sort(table(words), decreasing = TRUE))

  plot_data <- ref %>%
    mutate(index1 = match(.data[[names[1]]], ordered_words),
           index2 = match(.data[[names[2]]], ordered_words)) %>%
    mutate(slope = log10(index1) - log10(index2)) %>%
    pivot_longer(index1:index2) %>%
    mutate(value = log10(value),
           label = ifelse(name == "index1", 
                          .data[[names[1]]], 
                          .data[[names[2]]]),
           name = factor(name, c("index1", "index2"), names))
  
  limit <- max(abs(plot_data$slope)) * c(-1, 1)

  plot_data %>%
    ggplot(aes(name, value, group = .data[[names[2]]], label = label)) +
    geom_line(aes(color = slope), size = 1) +
    scale_y_reverse(labels = function(x) round(10 ^ x)) +
    geom_text_repel(data = subset(plot_data, name == names[1]),
                    aes(segment.color = slope),
                    nudge_x       = -0.1,
                    segment.size  = 1,
                    direction     = "y",
                    hjust         = 1) + 
    geom_text_repel(data = subset(plot_data, name == names[2]),
                    aes(segment.color = slope),
                    nudge_x       = 0.1,
                    segment.size  = 1,
                    direction     = "y",
                    hjust         = 0) + 
    scale_color_paletteer_c(palette, 
                            limit = limit,
                            aesthetics = c("color", "segment.color"), 
                            ...) +
    guides(color = "none", segment.color = "none") +
    theme_minimal() +
    theme(panel.border = element_blank(), 
          panel.grid.major.x = element_blank(), axis.text.x = element_text(size = 15)) +
    labs(x = NULL, y = "Word Rank")
}
```

Now we can recreate the previous chart with ease

```{r}
ref <- tribble(
  ~Men, ~Women,
  "he", "she",
  "his", "her",
  "man", "woman",
  "men", "women",
  "boy", "girl",
  "he's", "she's",
  "he'd", "she'd",
  "he'll", "she'll",
  "himself", "herself"
)

words <- hcandersen_en %>% 
  unnest_tokens(word, text) %>%
  pull(word)

plot_fun(words, ref, direction = -1) +
  labs(title = "Masculine gendered words appeared more often in H.C. Andersen's fairy tales")
```

## Gallery `r emo::ji("framed_picture")`

```{r}
ref <- tribble(
  ~Men, ~Women,
  "he", "she",
  "his", "her",
  "man", "woman",
  "men", "women",
  "boy", "girl",
  "himself", "herself"
)

words <- janeaustenr::austen_books() %>% 
  unnest_tokens(word, text) %>%
  pull(word)

plot_fun(words, ref, direction = -1) +
  labs(title = "Masculine gendered words appeared less often in Jane Austen Novels")
```

More examples using the [tidygutenbergr](https://github.com/EmilHvitfeldt/tidygutenbergr) package.

```{r, message=FALSE}
ref <- tribble(
  ~Men, ~Women,
  "he", "she",
  "his", "her",
  "man", "woman",
  "men", "women",
  "boy", "girl",
  "he's", "she's",
  "himself", "herself"
)

words <- tidygutenbergr::phantom_of_the_opera() %>% 
  unnest_tokens(word, text) %>%
  pull(word)

plot_fun(words, ref, "scico::berlin") +
  labs(title = "Masculine gendered words appeared more often in Phantom of the Opera")
```

```{r, message=FALSE}
ref <- tribble(
  ~Positive, ~Negative,
  "good", "bad",
  "pretty", "ugly",
  "friendly", "hostile"
)

words <- tidygutenbergr::dracula() %>% 
  unnest_tokens(word, text) %>%
  pull(word)

plot_fun(words, ref, palette = "scico::tokyo") +
  labs(title = "Positive adjectives appeared more often in Dracula")
```
