---
title: Changing Glyph in legend in ggplot2
date: '2019-06-17'
slug: changing-glyph-in-ggplot2
categories: [ggplot2]
image:
  preview_only: true
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(
  echo = TRUE,
  cache = TRUE,
  collapse = TRUE,
  fig.width = 7, 
  fig.align = 'center',
  fig.asp = 0.618, # 1 / phi
  out.width = "700px")
knit_hooks$set(optipng = hook_optipng)
opts_chunk$set("optipng" = "-o5")
```

# Introduction

The newest version of [ggplot2 3.2.0](https://www.tidyverse.org/articles/2019/06/ggplot2-3-2-0/) gave us the ability to change the glyph in the legend like so

```{r}
library(ggplot2)

ggplot(economics_long, aes(date, value01, colour = variable)) +
  geom_line(key_glyph = "timeseries")
```

And they can likewise be specified with the `draw_key_*` functions as well

```{r}
ggplot(economics_long, aes(date, value01, colour = variable)) +
  geom_line(key_glyph = draw_key_timeseries)
```

# Showcase

The following is all the available `draw_key_*` functions in ggplot2. Notice that the dark gray color in dotplot and polygon is a result of a unspecified `fill` aesthetic. Code to generate these figures can be found at the end of this post.

```{r echo=FALSE, fig.asp=0.6, fig.width=3, dpi=300}
library(dplyr)
library(magrittr)
library(ggplot2)
library(grid)

draws <- ls(getNamespace("ggplot2"), pattern = "^draw_key_")

legend_fun <- function(x) {
  ggg <- economics_long %>%
    mutate(variable = factor(variable, labels = paste("Option", LETTERS[1:5]))) %>%
    ggplot(aes(date, value01, colour = variable)) +
  geom_line(key_glyph = get(x)) +
    labs(color = x) 
  
  legend <- cowplot::get_legend(ggg)
  
  grid.newpage()
  grid.draw(legend)
}

purrr::walk(draws[1:12], legend_fun)
p <- ggplot(mtcars, aes(wt, mpg, label = rownames(mtcars))) + 
  geom_text(aes(colour = factor(ceiling(seq_len(nrow(mtcars)) %% 5), labels = paste("Option", LETTERS[1:5])))) +
  labs(color = "draw_key_text")
legend <- cowplot::get_legend(p)

grid.newpage()
grid.draw(legend)
purrr::walk(draws[14:16], legend_fun)
```

# Custom glyph key

Since the `draw_key_*` function just return a grob, you can use spend some time and create your own custom glyphs! Taking inspiration from `draw_key_boxplot`

```{r}
draw_key_boxplot
```

will I try to make a glyph by myself using both points and lines.

```{r, message=FALSE}
library(grid)
library(rlang)
draw_key_smile <- function(data, params, size) {
  grobTree(
    pointsGrob(0.25, 0.75, size = unit(.25, "npc"), pch = 16),
    pointsGrob(0.75, 0.75, size = unit(.25, "npc"), pch = 16),
    linesGrob(c(0.9, 0.87, 0.78, 0.65, 0.5, 0.35, 0.22, 0.13, 0.1), 
              c(0.5, 0.35, 0.22, 0.13, 0.1, 0.13, 0.22, 0.35, 0.5)),
    gp = gpar(
      col = data$colour %||% "grey20",
      fill = alpha(data$fill %||% "white", data$alpha),
      lwd = (data$size %||% 0.5) * .pt,
      lty = data$linetype %||% 1
    )
  )
}

ggplot(economics_long, aes(date, value01, colour = variable)) +
  geom_line(key_glyph = draw_key_smile)
```

And it looks so happy!

# Appendix

```{r, eval=FALSE}
library(dplyr)
library(magrittr)
library(ggplot2)
library(grid)

draws <- ls(getNamespace("ggplot2"), pattern = "^draw_key_")

legend_fun <- function(x) {
  ggg <- economics_long %>%
    mutate(variable = factor(variable, labels = paste("Option", LETTERS[1:5]))) %>%
    ggplot(aes(date, value01, colour = variable)) +
  geom_line(key_glyph = get(x)) +
    labs(color = x) 
  
  legend <- cowplot::get_legend(ggg)
  
  grid.newpage()
  grid.draw(legend)
}

purrr::walk(draws[1:12], legend_fun)
p <- ggplot(mtcars, aes(wt, mpg, label = rownames(mtcars))) + 
  geom_text(aes(colour = factor(ceiling(seq_len(nrow(mtcars)) %% 5), labels = paste("Option", LETTERS[1:5])))) +
  labs(color = "draw_key_text")
legend <- cowplot::get_legend(p)

grid.newpage()
grid.draw(legend)
purrr::walk(draws[14:16], legend_fun)
```

```{r details, echo=FALSE}
library(details) 

sessioninfo::session_info() %>%
  details::details(summary = 'session information')
```
