---
title: Use prismatic with after_scale() for finer control of colors in ggplot2
date: '2020-02-25'
categories: [color, ggplot2]
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(
  echo = TRUE,
  cache = TRUE,
  collapse = TRUE,
  fig.width = 7, 
  fig.align = 'center',
  fig.asp = 0.618, # 1 / phi
  out.width = "700px")
knit_hooks$set(optipng = hook_optipng)
opts_chunk$set("optipng" = "-o5")
```

With the release of version 3.3.0 of ggplot2 came the ability to have more control over the aesthetic evaluation.
This allows us to modify the colors of the mapped palettes with [prismatic](https://github.com/EmilHvitfeldt/prismatic) now easier than ever.

## Packages `r emo::ji("package")`

We load the essential packages to wrangle, collect data (we will use tweets), scrape websites and handle emojis.

```{r packages, message=FALSE}
library(ggplot2)
library(prismatic)
```

## Examples

Suppose you have a simple bar chart and you have added colors to each bar.

```{r}
ggplot(diamonds, aes(cut)) +
  geom_bar(aes(fill = cut))
```

Next, suppose you would like to add a border around each bar.
Traditionally you could add a single color like black but it isn't that satisfying as it doesn't have any relation to the mapped colors.

```{r}
ggplot(diamonds, aes(cut)) +
  geom_bar(aes(fill = cut), color = "black")
```

now that `after_scale()` is available for us we can base the color based on the mapped fill colors.
Below I have used `clr_darken()` to create a border that is just slightly darker than the fill color.

```{r}
ggplot(diamonds, aes(cut)) +
  geom_bar(aes(fill = cut, color = after_scale(clr_darken(fill, 0.3))))
```

this could also have been done in reverse by supplying the color and modifying the fill after.
Notice how we are able to chain multiple color modifications together.
Here we hare taking the color, then desaturating it followed by some lighting.

```{r}
ggplot(diamonds, aes(cut)) +
  geom_bar(aes(color = cut, 
               fill = after_scale(clr_lighten(clr_desaturate(color), 
                                              space = "combined"))))
```

If you only need to specify one color directly you can use the `stage()` function.

```{r}
ggplot(diamonds, aes(cut)) +
  geom_bar(aes(fill = stage(start = cut, 
                            after_scale = clr_lighten(fill, space = "combined"))))
```

```{r details, echo=FALSE}
library(details)

sessioninfo::session_info()  %>%
  details::details(summary = 'current session info')
```
