---
title: Real Emojis in ggplot2
date: '2020-01-02'
categories: [ggplot2]
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(
  echo = TRUE,
  cache = TRUE,
  collapse = TRUE,
  fig.width = 7, 
  fig.align = 'center',
  fig.asp = 0.618, # 1 / phi
  out.width = "700px")
knit_hooks$set(optipng = hook_optipng)
opts_chunk$set("optipng" = "-o5")
```

I have been trying to use [Emojis](https://en.wikipedia.org/wiki/Emoji) for a long time. 
It was actually part of my very first [post](https://www.hvitfeldt.me/blog/2017-world-press-freedom-index-with-emojis/) on this blog.
Others have made progress such as with [emojifont](https://cran.r-project.org/web/packages/emojifont/vignettes/emojifont.html), but it is not using the classical [Apple Color Emoji](https://en.wikipedia.org/wiki/Apple_Color_Emoji) font which is the most commonly recognized.
I made a breakthrough when I was writing the [packagecalander](https://www.hvitfeldt.me/packagecalendar/2019/) entry on [ggtext](https://github.com/clauswilke/ggtext).
While the method is the best I have found it does have some cons.

Pros:

- Actually works
- Doesn't require the use of SVG
- Previews nicely

Cons:

- Relies on experimental package **ggtext**
- Needs web scraping
- Required access to the internet to render
- Size can't be adjusted using the size aesthetic 

All in all, it is a fair trade for my needs.

## Packages `r emo::ji("package")`

We load the essential packages to wrangle, collect data (we will use tweets), scrape websites and handle emojis.

```{r packages, message=FALSE}
library(tidyverse)
library(rtweet)
library(rvest)
# devtools::install_github("clauswilke/ggtext")
library(ggtext)
library(emo)
```

## Getting the tweets `r emo::ji("bird")`

For a simple dataset where we find emojis I'm going to get some tweets with the word "happy".

```{r, eval=FALSE}
happy <- search_tweets("happy", include_rts = FALSE, n = 1000)
```

```{r tweets, echo=FALSE, message=FALSE}
happy <- search_tweets("happy", include_rts = FALSE, n = 1000, retryonratelimit = TRUE)
```

we can use the `ji_extract_all()` function from the [emo](https://github.com/hadley/emo) package.
This will give us a list of emojis so we can use the `unnest()` function to get back to a tidy format.
I'm going to do a simple `count()` of the emojis for the following visualizations. 

```{r top_happy}
happy_emojis <- happy %>%
  mutate(emoji = emo::ji_extract_all(text)) %>%
  unnest(cols = c(emoji)) %>%
  count(emoji, sort = TRUE)
```

Next is where the magic happens.
We don't have a way to displays emojis in **ggplot2**, but we can use **ggtext** to embed images into the text using HTML.
Now we just need to get an image of each emoji.
The following function will accept an emoji as a string and return the URL to a .png of that emoji.

```{r emoji_to_link}
emoji_to_link <- function(x) {
  paste0("https://emojipedia.org/emoji/",x) %>%
    read_html() %>%
    html_nodes("tr td a") %>%
    .[1] %>%
    html_attr("href") %>%
    paste0("https://emojipedia.org/", .) %>%
    read_html() %>%
    html_node('div[class="vendor-image"] img') %>%
    html_attr("src")
}
```

Then this function will take that URL and construct the necessary HTML code to show the emoji PNGs.

```{r link_to_img}
link_to_img <- function(x, size = 25) {
  paste0("<img src='", x, "' width='", size, "'/>")
}
```

To be courteous we are only going to scrape the emojis we are actually going to use.
So we will `slice()` the 10 most frequent emojis.
We will also be adding a 5 second delay using `slowly()` and `rate_delay()` from **purrr**.

```{r top_happy2, dependson="link_to_img"}
top_happy <- happy_emojis %>%
  slice(1:10) %>%
  mutate(url = map_chr(emoji, slowly(~emoji_to_link(.x), rate_delay(1))),
         label = link_to_img(url))
```

## emoji-scatter plot `r emo::ji("chart_with_upwards_trend")`

Now we can use the `geom_richtext()` function from **ggtext** to create a emoji scatter chart.

```{r plot1, dependson="top_happy2"}
top_happy %>%
  ggplot(aes(emoji, n, label = label)) +
  geom_richtext(aes(y = n), fill = NA, label.color = NA, # remove background and outline
                label.padding = grid::unit(rep(0, 4), "pt") # remove padding
  ) +
  theme_minimal()
```

This is a little off, so lets other these by counts and put them over a bar chart.
I'm also going to the x-axis ticks and text.

```{r plot2, warning=FALSE, dependson="top_happy2"}
offset <- max(top_happy$n) / 20

top_happy %>%
  ggplot(aes(fct_reorder(emoji, n, .desc = TRUE), n, label = label)) +
  geom_col() +
  geom_richtext(aes(y = n + offset), fill = NA, label.color = NA,
                label.padding = grid::unit(rep(0, 4), "pt")
  ) +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank()) +
  labs(x = NULL) +
  theme_minimal()
```

## Emojis in labels and text `r emo::ji("bar_chart")`

We are not only limited to using emojis in the geoms.
We can set the text element using emojis to `element_markdown()`.
Below we have the same bar chart as above but with the emoji as labels below instead of on top.

```{r plot3, warning=FALSE, dependson="top_happy2"}
top_happy %>%
  ggplot(aes(fct_reorder(label, n, .desc = TRUE), n)) +
  geom_col() +
  theme_minimal() +
  theme(axis.text.x = element_markdown()) +
  labs(x = NULL)
```

## Adding a splash of color `r emo::ji("rainbow")`

We can employ a little more scraping and color calculations to had colors to the bars according to the colors of the emoji.
The following function takes a URL to a .png file and returns the most common color that isn't pure black or pure white.

```{r mean_emoji_color}
mean_emoji_color <- function(x) {
  data <- png::readPNG(RCurl::getURLContent(x))
  color_freq <- names(sort(table(rgb(data[,,1], data[,,2], data[,,3])), 
                           decreasing = TRUE))
  setdiff(color_freq, c("#FFFFFF", "#000000"))[1]
}
```

We apply this to all the emoji URLs and color the bars accordingly. 

```{r plotdata, warning=FALSE, dependson=c("top_happy2", "packages")}
plot_data <- top_happy %>%
  mutate(color = map_chr(url, slowly(~mean_emoji_color(.x), rate_delay(1))))

plot_data %>%
  ggplot(aes(fct_reorder(label, n, .desc = TRUE), 
             color = color, 
             fill = unclass(prismatic::clr_lighten(color, 0.4)), n)) +
  geom_col() +
  scale_fill_identity() +
  scale_color_identity() +
  theme_minimal() +
  theme(axis.text.x = element_markdown()) +
  labs(x = NULL, y = "Count",
       title = "Emojis used in (small sample) of 'happy' tweets",
       subtitle = "Displayed in ggplot2!!!",
       caption = "@Emil_Hvitfeldt")
```

## Final note `r emo::ji("spiral_notepad")`

If you want to use emojis in the text you need to call `theme_*()` before `theme()` such that `element_markdown()` isn't being overwritten.

```{r details, echo=FALSE}
library(details)

sessioninfo::session_info()  %>%
  details::details(summary = 'current session info')
```

